untitled

### 2016-04-23 model.tex
update

\subsection{Vertical shear}
The mixture possess vertical shear. To see this, we eliminate $\Phi$
between Eq. \ref{steady_momr}---\ref{steady_momz} to 
obtain 
\begin{align}\label{vshear}
  r\frac{\p \Omega^2}{\p z} &= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
  & = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
    P}{\p z}\frac{\p s}{\p r} \right),\notag
\end{align}
where
\begin{align}
   s \equiv \ln \frac{P}{\rho}
\end{align}
may be consider as a pseudo-entropy for the mixture.

@@

\subsubsection{Order-of-magnitude estimates}
Let us assume the gas and dust distributions have a Gaussian
distribution vertically,
\begin{align}
  \rho_\mathrm{g,d}(r,z) = \rho_\mathrm{g0,d0}(r) \exp
      {\left[-\frac{z^2}{2H^2_\mathrm{g,d}(r)}\right]}.
\end{align}
Then the dust-to-gas ratio is 
\begin{align}
  \tepsilon = \tepsilon_0(r)
  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}, 
\end{align}
with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
\begin{align}
  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
  \frac{1}{H_\mathrm{g}^2}. 
\end{align}
%We assume the dust layer is much thinner than the gas scale-height,
%$H_\mathrm{d}\ll H_\mathrm{g}$, so that 
%\begin{align}
% \widetilde{H} \sim H_\mathrm{d}. 
%\end{align}

We further assume
%the gas and dust
%distributions are smoothly varying in the radial direction so that  
%\begin{align}
%  \frac{\p}{\p r}\sim \frac{p}{r}, 
%\end{align}
the disk is roughly Keplerian so that
\begin{align}
  \kappa \sim \OmK, 
\end{align}
and the gas sound-speed is 
\begin{align}
  c_s \sim h_\mathrm{g}r\OmK, 
\end{align}
where $h_\mathrm{g}\ll1$ is the aspect-ratio of the gas component.

%Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
%dust-to-gas ratio $\tepsilon$,           
%we find 
%\begin{align}\label{ineq_est}
%  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
%  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
%  \left[
%    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
%    \right]>0
%\end{align}
%is required \emph{for stability.} 
%This inequality is easily satisifed in typical disk models where 
%the dust layer is much thinner than the gas scale-height, so
%that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
%Then the magnitude the only destabilizing term (the third term) is negligible. According to Eq. \ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.

\begin{align} 
  &1 + \frac{\tepsilon}{\left(1+
    \tepsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
  \frac{z^2}{\widetilde{H}^2}\right)>0,   \\ 
&1 - \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right) > 0,
\end{align}

>>>>

\subsection{Vertical shear}
The mixture possess vertical shear. To see this, we eliminate $\Phi$
between Eq. \ref{steady_momr}---\ref{steady_momz} to 
obtain 
\begin{align}\label{vshear}
  r\frac{\p \Omega^2}{\p z} &= \frac{\p\ln{\rho}}{\p r}\frac{\p}{\p
    z}\left[c_s^2(1-\epsilon)\right] - \frac{\p\ln{\rho}}{\p z}
  \frac{\p}{\p r} \left[c_s^2(1-\epsilon)\right]\\  
  & = \frac{1}{\rho}\left(\frac{\p P}{\p r}\frac{\p s}{\p z} -\frac{\p
    P}{\p z}\frac{\p s}{\p r} \right),\notag
\end{align}
where
\begin{align}
   s \equiv \ln \frac{P}{\rho}
\end{align}
is a pseudo-entropy for the mixture.

@@

\subsubsection{Order-of-magnitude estimates}
We evaluate the above stability criteria by assuming the gas and dust
distributions have a Gaussian 
distribution vertically, 
\begin{align}
  \rho_\mathrm{g,d}(r,z) = \rho_\mathrm{g0,d0}(r) \exp
      {\left[-\frac{z^2}{2H^2_\mathrm{g,d}(r)}\right]}.
\end{align}
We note that $H_\mathrm{g}=h_gr$ by definition. The dust-to-gas ratio
is  
\begin{align}
  \tepsilon = \tepsilon_0(r)
  \exp{\left[-\frac{z^2}{2\widetilde{H}^2(r)}\right]}, 
\end{align}
with $\tepsilon_0= \rho_\mathrm{d,0}/\rho_{g_0}$, and  
\begin{align}
  \frac{1}{\widetilde{H}^2} \equiv \frac{1}{H_\mathrm{d}^2} -
  \frac{1}{H_\mathrm{g}^2}. 
\end{align}
We further assume the disk is approximately Keplerian so that
\begin{align}
  \kappa \simeq \Omega. 
\end{align}
%Inserting the above approximations into Eq. \ref{criterion2} and writing the result in terms of the
%dust-to-gas ratio $\tepsilon$,           
%we find 
%\begin{align}\label{ineq_est}
%  1 + \frac{\tepsilon h_\mathrm{g}^2}{\left(1 + \tepsilon\right)^2}
%  \left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right| 
%  \left[
%    \left|\frac{\p\ln{\rhog}}{\p\ln{r}}\right| - \frac{\widetilde{H}^2}{H^2_\mathrm{g}}\left|\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right|  
%    \right]>0
%\end{align}
%is required \emph{for stability.} 
%This inequality is easily satisifed in typical disk models where 
%the dust layer is much thinner than the gas scale-height, so
%that $\widetilde{H}\sim H_\mathrm{d}\ll H_\mathrm{g}$, and where the radial profiles smoothly vary on a scale of $r$.
%Then the magnitude the only destabilizing term (the third term) is negligible. According to Eq. \ref{ineq_est}, instability is on%ly possible if there is a rapid variation the radial profile of the dust-to-gas ratio.   
Then the above criteria becomes
\begin{align}
  &1 + \frac{\tepsilon}{\left(1+
    \tepsilon\right)^2}\left(h_\mathrm{g}^2\frac{\p\ln{\rhog}}{\p\ln{r}}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} + 
  \frac{z^2}{\widetilde{H}^2}\right)>0 \label{stability_est1},   \\ 
&1 - \frac{\tepsilon
  h_\mathrm{g}^2}{\left(1+\tepsilon\right)^2}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}}\left(-\frac{\p\ln{\rhog}}{\p\ln{r}}+\frac{\widetilde{H}^2}{H_\mathrm{g}^2}\frac{\p\ln{\tepsilon}}{\p\ln{r}}\right)
  > 0 \label{stability_est2},
\end{align}
\emph{for stability}.

Eq. \ref{stability_est1}---\ref{stability_est2} are easily satisfied
in typical accretion disk models where radial gradients are $O(1/r)$ and
$h_\mathrm{g},\tepsilon, \widetilde{H}/H_\mathrm{g}\ll 1$, since then the
magnitude of the second term on the left-hand-side  of either
inequality is much less than unity. Such disks are stable against
axisymmetric perturbations.

\subsubsection{Possible instability at dust edges}
Notice the left-hand-side of Eq. \ref{stability_est2} is a quadratic in
$\p_r\tepsilon$. It is possible to
violate this inequality for sufficiently large (in magnitude) radial
gradients  in the dust-to-gas ratio, 
\begin{align}
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} > S_+ \quad \mathrm{or} \quad 
  \frac{\p\ln{\tepsilon}}{\p\ln{r}} < S_-,
\end{align}
\emph{for instability}, where
\begin{align}\label{spm}
S_\pm = \frac{1}{2}\frac{H_\mathrm{g}^2}{\widetilde{H}^2} 
  \left[
  \frac{\p\ln{\rhog}}{\p\ln{r}} \pm 
  \sqrt{
  \left(\frac{\p\ln{\rhog}}{\p\ln{r}}\right)^2 + 
  4 \frac{\widetilde{H}^2}{H_\mathrm{g}^2}
  \frac{\left(1+\tepsilon\right)^2}{\tepsilon h_\mathrm{g}^2}
  }
  \,\right]. 
\end{align}

Now, if $\p_r\rhog<0$, as typical in
accretion models, then it is easier to achieve instability at a given
radius for rapidly increasing dust-to-gas ratios ($\p_r\tepsilon >
0$), and vice versa. 
%This might occur, for example, when dust is
%inserted in a finite radial range within a wider gaseous disk. 
%Physically, one might expect
%$\p_r\rhog\p_r\tepsilon > 0$ because

\emph{If} the $\p_r\rhog$ can be neglected in Eq. \ref{spm},
then instability requires
\begin{align}
  \left|\frac{\p\ln{\tepsilon}}{\p r}\right| \gtrsim
  \frac{1}{\widetilde{H}}\frac{\left(1+\tepsilon\right)}{\sqrt{\tepsilon}}. 
\end{align}
That is, if the dust-to-gas ratio varies on a lengthscale less than
$O(\widetilde{H}\sqrt{\tepsilon})$ for $\tepsilon\ll 1$, or less than
$O(\widetilde{H}/\sqrt{\tepsilon})$ for $\tepsilon\gg 1$, then the system is
potentially unstable.
